---
title: "Parkinson_Linear_Regression"
author: "Yemi Adekoluejo"
date: "2024-02-12"
output: html_document
---
Reading in the 'Parkinsons_updrs' dataset and names:
```{r}
PUD <- read.csv('parkinsons_updrs.data1')
```
Checking for any Null Values in the dataset. Obtaining summary information about the dataset.
```{r}
View(PUD)
sum(is.na(PUD))
summary(PUD)
```
Removing 'total_UPDRS' as this will be the variable the linear regression model will be predicting.
```{r}
PUD = subset(PUD, select = -c(total_UPDRS))
```
Determining the correlation coefficient:
```{r}
cor(PUD)
```
```{r}
set.seed(2) #set seeds guarantees results are reproducible 
total_examples <- nrow(PUD)
```
Creating training set with 80% of the samples and test set with 20%:
```{r}
train_size <- round(0.8 * total_examples) 
test_size <- total_examples - train_size
```
Randomizing training set and test set:
```{r}
shuffle <- sample(1:total_examples, total_examples) 
train_shuffle <- shuffle[1:train_size]
test_shuffle <- shuffle[(train_size + 1):total_examples]
```
Viewing preprocessed training data and testing data:
```{r}
training_data <- PUD[train_shuffle, ]
testing_data <- PUD[test_shuffle, ]
View(training_data)
print(( nrow(training_data)))
print(( nrow(testing_data)))
```
First Linear Regression Model using all associated variables to determing which vairables are the most influential.
Viewing R^2 and Residual standard error to evaluate performance. High R^2 and low RSE provide evidence for a good accurate model. The most significant variables are Shimmer.APQ11, NHR, HNR, DFA, PPE, subject, age, sex, test_time, Jitter.abs
```{r}
lm.PUD <- lm(motor_UPDRS~., data = PUD)
summary(lm.PUD)
summary(lm.PUD)$r.sq 
summary(lm.PUD)$sigma
```
Using an Interaction term on the models most influential vairable and evaluating changes to models performance
```{r}
lm.PUDM <- lm(motor_UPDRS ~ Shimmer.APQ11*., data=PUD)
summary(lm.PUDM)
```
Adding non-linear transformation of predictors and evaluating model performance.
```{r}
lm.PUDI <- lm(motor_UPDRS~I(Shimmer.APQ11)+NHR+HNR+DFA+PPE+subject.+age+sex+test_time, data = PUD)
summary(lm.PUDI)
```
Diagnostic plots of each iteration of the linear regression model
```{r}
plot(lm.PUD)
plot(lm.PUDI)
plot(lm.PUDM)
```
After analyzing the diagnostic plots of each linear regression model created although all are similar in terms of results the lm.PUDI (linear regression model with interaction term) had the largest spread of data in the residuals vs fitted plot while maintaining an uncorrelated structure also when analyzing the residuals vs leverage plot there are more less high leverage data points that stray away from the normal line in lm.PUDI

